---
title: "CH 1"
author: "Dan Schumacher"
format: html
editor: visual
---

```{r}
#install.packages('palmerpenguins')
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

# R For Data Science

## Chapter 2 Code Along

We will be using penguins

```{r}
?penguins
penguins
glimpse(penguins)
```

```{r}

```

```{r}
#creates a blank plot
ggplot(data = penguins)

#with mapping we add the axises
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
)

#Add Geom_SOMETHING to plot data
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point()

#color by species
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm,
                y = body_mass_g,
                color = species)
) +
  geom_point()

#add trend lines
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm,
                y = body_mass_g,
                color = species)
) +
  geom_point() +
  geom_smooth(method = 'lm')

#Since we want points to be colored based on species but don’t want the lines to be separated out for them, we should specify color = species for geom_point()
#lets also add shape aesthetic
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm,
                y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = 'lm')

#improve our labels
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm,
                y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = 'lm') +

labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  ) +
  scale_color_colorblind()
```

## **2.2.5 Exercises**

1.  How many rows are in `penguins`? How many columns?

    ```{r}
    paste('rows: ', nrow(penguins), 'cols: ', ncol(penguins))
    ```

2.  What does the `bill_depth_mm` variable in the `penguins` data frame describe? Read the help for [`?penguins`](https://allisonhorst.github.io/palmerpenguins/reference/penguins.html) to find out.

    "a number denoting bill depth (millimeters)"

    ```{r}
    ?penguins
    ```

3.  Make a scatterplot of `bill_deptph_mm` vs. `bill_length_mm`. That is, make a scatterplot with `bill_depth_mm` on the y-axis and `bill_length_mm` on the x-axis. Describe the relationship between these two variables.

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(x = bill_length_mm, y = bill_depth_mm)
    ) +
      geom_point()
    ```

4.  What happens if you make a scatterplot of `species` vs. `bill_depth_mm`? What might be a better choice of geom?

    a\. You get a clear idea of the range of bill for each species

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(x = species, y = bill_depth_mm)
    ) +
      geom_point()
    ```

5.  Why does the following give an error and how would you fix it?

    ```{r}
    ggplot(data = penguins) + 
      geom_point()
    ```

    It gives an error because mapping = aes() have not been defined. R doesn't know what to plot on the x or y axis. To fix it you need to provide that info.

6.  What does the `na.rm` argument do in [`geom_point()`](https://ggplot2.tidyverse.org/reference/geom_point.html)? What is the default value of the argument? Create a scatterplot where you successfully use this argument set to `TRUE`.

    na.rm argument removes missing values. It is set to false by default.

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(x = bill_length_mm, y = bill_depth_mm)
    ) +
      geom_point(na.rm = T)
    ```

7.  Add the following caption to the plot you made in the previous exercise: "Data come from the palmerpenguins package." Hint: Take a look at the documentation for [`labs()`](https://ggplot2.tidyverse.org/reference/labs.html).

    ```{r}
    {r}
    ggplot(
      data = penguins,
      mapping = aes(x = bill_length_mm, y = bill_depth_mm)
    ) +
      geom_point(na.rm = T) +
      labs(
        caption = "Data come from the palmerpenguins package."
      )
    ```

8.  Recreate the following visualization. What aesthetic should `bill_depth_mm` be mapped to? And should it be mapped at the global level or at the geom level?

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(x = flipper_length_mm, y = body_mass_g)
    ) +
      geom_point(mapping = aes(color = bill_depth_mm)) +
      geom_smooth()
    ```

9.  Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
    ) +
      geom_point() +
      geom_smooth(se = FALSE)
    ```

10. Will these two graphs look different? Why/why not?

    No. the aesthetics are just applied in different layers but are the same

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(x = flipper_length_mm, y = body_mass_g)
    ) +
      geom_point() +
      geom_smooth()

    ggplot() +
      geom_point(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
      ) +
      geom_smooth(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
      )
    ```

## Chapter 2 Code Along

As long as our args are in the right order, we don't need to explicitly state them.

```{r}
ggplot(penguins, 
       aes(x = flipper_length_mm,
           y = body_mass_g)) + 
  geom_point()

#even cleaner with pipe operator
penguins |> 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point()
```

```{r}
#bar plots
ggplot(penguins, aes(x = species)) +
  geom_bar()

#reodered
ggplot(penguins, aes(x = fct_infreq(species))) +
  geom_bar()
```

## **2.4.3 Exercises**

1.  Make a bar plot of `species` of `penguins`, where you assign `species` to the `y` aesthetic. How is this plot different?

    *It is sideways*

    ```{r}
    ggplot(
      penguins,
      aes(y = species)
    ) +
      geom_bar()
    ```

2.  How are the following two plots different? Which aesthetic, `color` or `fill`, is more useful for changing the color of bars?

    One's outline is red while the other's whole shape is red

    ```{r}
    ggplot(penguins, aes(x = species)) +
      geom_bar(color = "red")

    ggplot(penguins, aes(x = species)) +
      geom_bar(fill = "red")
    ```

3.  What does the `bins` argument in [`geom_histogram()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html) do?

    It determines the number of bars on your histogram. How many categories to divide the numerical data into.

4.  Make a histogram of the `carat` variable in the `diamonds` dataset that is available when you load the tidyverse package. Experiment with different binwidths. What binwidth reveals the most interesting patterns?

    ```{r}
    ggplot(
      diamonds,
      aes(x=carat)
    ) +
      geom_histogram(binwidth = .25)
    ```

## 2.5.2 Exercises

1.  The `mpg` data frame that is bundled with the ggplot2 package contains 234 observations collected by the US Environmental Protection Agency on 38 car models. Which variables in `mpg` are categorical? Which variables are numerical? (Hint: Type [`?mpg`](https://ggplot2.tidyverse.org/reference/mpg.html) to read the documentation for the dataset.) How can you see this information when you run `mpg`?

    Cat: manufacturer model, trans, drv, fl, class

    Num: displ, year, cyl, cty hwy

    You can see this info right under the variable name on the tibble

    ```{r}
    mpg
    ```

2.  Make a scatterplot of `hwy` vs. `displ` using the `mpg` data frame. Next, map a third, numerical variable to `color`, then `size`, then both `color` and `size`, then `shape`. How do these aesthetics behave differently for categorical vs. numerical variables?

    *can't do numerical variable and shape. Not a continues thing to adjust.*

    ```{r}
    ggplot(
      mpg,
      aes(x =hwy, y = displ, size = year, color = year)
    ) +
      geom_point()
    ```

3.  In the scatterplot of `hwy` vs. `displ`, what happens if you map a third variable to `linewidth`?

    *I don't get it...*

    ```{r}
    ggplot(
      mpg,
      aes(x =hwy, y = displ,linewidth = year)
    ) +
      geom_point()
    ```

4.  What happens if you map the same variable to multiple aesthetics?

    *I don't get it...*

5.  Make a scatterplot of `bill_depth_mm` vs. `bill_length_mm` and color the points by `species`. What does adding coloring by species reveal about the relationship between these two variables? What about faceting by `species`?

    ```{r}
    ggplot(
      penguins,
      aes(x = bill_depth_mm,
             y = bill_length_mm,
             color = species)
    ) +
      geom_point()

    ggplot(
      penguins,
      aes(x = bill_depth_mm,
             y = bill_length_mm,
             color = species)
    ) +
      geom_point() + 
      facet_wrap(~species)
    ```

6.  Why does the following yield two separate legends? How would you fix it to combine the two legends?

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(
        x = bill_length_mm, y = bill_depth_mm, 
        color = species, shape = species
      )
    ) +
      geom_point() +
      labs(color = "Species")


    #get rid of the labels argument.
    ggplot(
      data = penguins,
      mapping = aes(
        x = bill_length_mm, y = bill_depth_mm, 
        color = species, shape = species
      )
    ) +
      geom_point()
    ```

7.  Create the two following stacked bar plots. Which question can you answer with the first one? Which question can you answer with the second one?

    1.  Dream Island has roughly 50% chinstrap to adelie species

    2.  adelie can be found on every island.

    ```{r}
    ggplot(penguins, aes(x = island, fill = species)) +
      geom_bar(position = "fill")

    ggplot(penguins, aes(x = species, fill = island)) +
      geom_bar(position = "fill")
    ```
