---
title: "CH 3"
format: html
editor: visual
---

# Chapter 3

## 3.5 Exercises

1.  Why does this code not work?

    ı != i

    Idk what that character is but it isn't an 'i'

```{r}
my_variable <- 10
my_varıable
```

2.  Tweak each of the following R commands so that they run correctly:

```{r}
#original
libary(todyverse)

ggplot(dTA = mpg) + 
  geom_point(maping = aes(x = displ y = hwy)) +
  geom_smooth(method = "lm)
```

```{r}
#updated
library(tidyverse)

ggplot(data = mpg,
       aes(x = displ, y = hwy)) +  
  geom_point() +
  geom_smooth(method = "lm")
```

3.  Press Alt + Shift + K. What happens? How can you get to the same place using the menus?

    ITS A SHORTCUT LIST!

4.  Let\'s revisit an exercise from the Section 2.6. Run the following lines of code. Which of the two plots is saved as `mpg-plot.png`? Why?

    ```{r}
    my_bar_plot <- ggplot(mpg, aes(x = class)) +
      geom_bar()
    my_scatter_plot <- ggplot(mpg, aes(x = cty, y = hwy)) +
      geom_point()
    ggsave(filename = "mpg-plot.png", plot = my_bar_plot)
    ```

The bar plot is the first to show! That is because it is the first one listed.
